
# "Reproducible Research: Peer Assessment 1"

## Load required libraries
```{r loadlibrary}
library(ggplot2)
library(lattice)
```

## Loading and preprocessing the data
```{r loadData}
file="C:/Users/Nicholas/Downloads/repdata%2Fdata%2Factivity/activity.csv"
activity<-read.csv(file, header = TRUE, colClass=c('integer', 'Date', 'integer'))
```

## Learn about the data.
```{r learn}
str(activity)
head(activity)
summary(activity)
```

## What is mean total number of steps taken per day?
In this section we will sum up the steps by day using the aggregate function.

daytotal<-aggregate(steps ~ date, activity, sum)

Next we will plot the daily step count on a histogram.

barplot(daytotal$steps,xlab="date",ylab="steps for day",main="Daily steps by date")

Next let us compare the mean vs the median of the total number of steps taken per day.

mean(daytotal$steps)
median(daytotal$steps)

## What is the average daily activity pattern?
When making this graph we will nolonger be looking at the daily amount and now in 5 minute intervals (the intervals given)

interval.total<-aggregate(steps ~ interval, activity, mean)
ggplot(interval.total,aes(interval,steps))+geom_line(color="blue",size=1)+labs(title = "Time Series Plot of the 5-minute Intervals", x = "5-minute intervals", y = "Average Number of Steps Taken")

## The 5-minute interval that, on average, contains the maximum number of steps 
max<-which.max(interval.total$steps)
interval.total$interval[max]

## Imputing missing values
The total number of rows with with NAs is.
sum(is.na(activity$steps))

##Fill in missing NA with the average number of steps in the same 5-min interval.
Create a new dataset (activity2) as the original and then replace the NA values. In this case we will be using the average for that interval.

imp <- activity # new dataset called imp
for (i in avg_step$interval) {
  imp[imp$interval == i & is.na(imp$steps), ]$steps <- 
    avg_step$steps[avg_step$interval == i]
}

#Check to see if the data has NA's anymore.

head(activity2) # no NAs 
  
##Make a histogram of the total number of steps taken each day and Calculate and report
##the mean and median total number of steps taken per day. Do these values differ from 
##the estimates from the first part of the assignment? What is the impact of imputing 
##missing data on the estimates of the total daily number of steps?

Keeping with the naming convention of activity2 being the second version of activity the follow convention will be used below.,

day.total2<-aggregate(steps ~ date, activity2, sum)
barplot(daytotal$steps,xlab="date",ylab="steps for day",main="Daily steps by date (NA's imputted)")

Next let us compare the mean vs the median of the total number of steps taken per day.

mean(day.total2$steps) # same as before.
median(day.total2$steps) # slightly higher than before


#Are there differences in activity patterns between weekdays and weekends?

#Create a new factor variable in the dataset with two levels – “weekday” and
#“weekend” indicating whether a given date is a weekday or weekend day.


activity2$day <- weekdays(activity2$date)
activity2$week <- ""
activity2[activity2$day == "Saturday" | activity2$day == "Sunday", ]$week <- "weekend"
activity2[!(activity2$day == "Saturday" | activity2$day == "Sunday"), ]$week <- "weekday"
activity2$week <- factor(activity2$week)

#Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
#and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
#See the README file in the GitHub repository to see an example of what this plot should look like using 
#simulated data.


averageacross <- aggregate(steps ~ interval + week, data = activity2, mean)

xyplot(steps ~ interval | week, data = avg_step_imp, type = "l", lwd = 2,
       layout = c(1, 2), 
       xlab = "5-minute intervals", 
       ylab = "Average Number of Steps Taken",
       main = "Average Number of Steps Taken at 5 Minute Intervals (weekdays vs weekend)")


